---
title: "Crowd4SDG Slack - Analysis for Report"
output: html_notebook
---

Analysis on the Crowd4SDG Slack included in the Report


```{r}

library("RColorBrewer") #Color Palette
library("wordcloud") #Wordcloud Library
library("jsonlite") #Handling JSON Files
library("ggplot2") #Plots
library("igraph") #Networks and Analysis
library("scales") #Plot Aesthetics
library("visNetwork") #D3 based Network Plots
library("plotly") #Plots
library("dplyr") #Dataframe Manipulations
library("qgraph") #Network analyis and Plots
library("networkD3") #Sankey

source("custom.R") #Some Custom Functions (courtesy Marc)

```

```{r}

users_df = read.csv("data_files/users_an.csv") #Users Meta

channels_df = read.csv("data_files/channels.csv") #Channels Meta
df_emoticon_responses = read.csv("data_files/emoticon_responses.csv")[,-c(1)] #Emoticon Responses
df_text_mentions = read.csv("data_files/text_mentions.csv")[,-c(1)] #Text Mentions
df_texts = read.csv("data_files/texts_channel.csv")[,-c(1)] #Messages
df_channel_join = read.csv("data_files/channel_join.csv")[,-c(1)] #Channel Join Messages
df_threads = read.csv("data_files/threads_channel.csv")[,-c(1)] #Threads

df_th_rep = read.csv("data_files/thread_replies.csv")[,c(-1)] #Thread Replies

teams = c("team-aquatics", "team-floodfinder", "team-toseetocare", "team-safewaterforhomeless", "team-wotter", "team-collaborativewatermonitoring", "teamfloodmaps", "team-filsteiner", "team-mnl", "team-rainwaterharvesting", "team-garudasavior", "team-safeskye", "team-fews", "team-rainwater-collection", "team-potamoi", "team-dailywater", "team_ecolution", "team-thousand-waters", "warbon-footprint")



```

Plot Functions

```{r}

#Toggle for arrowheads if directed

visPlot_ni = function(subgraph, communities = rep(1, length(V(subgraph))), nodesize = 10, edgewidth = 0, title="", textsize = nodesize, layout = "layout_nicely", directed = TRUE)
{
  nodes <- data.frame(id = V(subgraph)$name, name = V(subgraph)$print, group = communities)
  nodes$font.size<-textsize + 20
  nodes$font.color = "grey"
  nodes$font.face = "calibri"
  nodes$label = nodes$name
  nodes$value = nodesize
  edges <- data.frame(get.edgelist(subgraph))
  edges$width = edgewidth
  colnames(edges)<-c("from","to","width")
  plt = visNetwork(nodes, edges, height = "1000px", width = "100%",main = title)%>%
    visIgraphLayout(layout = layout) %>%
    visPhysics(solver = "repulsion", hierarchicalRepulsion = "nodeDistance") %>%
    visOptions(highlightNearest = TRUE) %>%
    visNodes(scaling = list(min = 10, max = 30)) %>%
    visEdges(smooth = T, color = list(color = "lightblue", highlight = "blue")) %>% 
    visGroups(groupname = "2", color = list(background = "lightgray",border = "black")) %>%
    visInteraction(keyboard = T,
                   dragNodes = T, 
                   dragView = T, 
                   zoomView = T)
  
  if (directed == TRUE)
    plt = plt %>% visEdges(arrows ="to")
  
  return(plt)
}

#Plot Function with qgraph library
plotNetwork <- function(Gsub, layout='default',maine='', coms=NULL, teams = NULL){
  
  
  if (is.null(coms)){
    com <- cluster_walktrap(Gsub)
    V(Gsub)$color <- com$membership+1
  } else {
    #col = merge(data.frame(team = coms), list)
    V(Gsub)$color <- coms
  }
  vertex.frame.color <- V(Gsub)$color
  
  size <- abs(strength(Gsub, loops=T))
  if (max(size, na.rm=T)>0){ 
    size <- 30 * size / max(size, na.rm=T)
  } else {
    size <- 1
  }
  
  if (is.null(E(Gsub)$weight)){
    width <- 1
    edge.color <- NA 
    l <- layout_with_fr
    
  } else{
    weight <- abs(E(Gsub)$weight)
    width <- .01 + 5 * weight / max(weight, na.rm=T)
    edge.color <- ifelse(E(Gsub)$weight > 0, 'gray', 'blue')
    e <- get.edgelist(Gsub, names=F)
    e <- cbind(e, E(Gsub)$weight)
    l <- qgraph.layout.fruchtermanreingold(e,
                                           vcount=vcount(Gsub),
                                           area=1*(vcount(Gsub)^2),
                                           repulse.rad=5*(vcount(Gsub)^2))
    
  }
  
  if (layout!='default'){
    l <- layout
  }
  
  plot(Gsub,
       layout=l,
       edge.width = width,
       edge.color=edge.color,
       vertex.label = V(Gsub)$Name,
       vertex.label.color = "black",
       vertex.label.cex = 0.7,
       vertex.size = 3 * sqrt(size),
       edge.curved=0.1,
       vertex.frame.color = vertex.frame.color, 
       edge.arrow.width = 0.1,
       edge.arrow.size = 0.1,
#       main=maine,
#       main.cex = 0.3,
       col='gray'
)
  title(main = maine , cex.main = 0.7)
  return(l)
}

```

# Sort Activity by Week

```{r}

Sys.setenv(TZ="GMT")
df_texts$source_message_timestamp = as.POSIXct(df_texts$source_message_timestamp, origin = "1970-01-01")

df_channel_activity_weekly = data.frame()

for (i in unique(df_texts$channel))
{
  if (T)
  {
    obj = hist(df_texts$source_message_timestamp[df_texts$channel == i], breaks = "weeks", plot = F)
    for (j in 1:length(obj$counts))
    {
      df_channel_activity_weekly = rbind(df_channel_activity_weekly, data.frame(channel = i, activity = obj$counts[j], weeks = obj$breaks[j]))
    }
  }
  
}
```

# Wordclouds

```{r,fig.height=4, fig.width=4}

list = sort(unique(df_channel_activity_weekly$weeks))
weeks = list
wc = list()

dir.create('viz', showWarnings = F)


for (i in list[1:(length(list))])
{
  set.seed(1234)
  # print(as.POSIXct(i, origin = "1970-01-01"))
  pdf(paste0('viz/wordclouds_week',match(i,list),'.pdf'), 10, 10)
  # par(mfrow=c(3,3))
   wc = wordcloud(words = df_channel_activity_weekly$channel[df_channel_activity_weekly$weeks == i], 
                  freq = df_channel_activity_weekly$activity[df_channel_activity_weekly$weeks == i], 
                  min.freq = 0,max.words=100, random.order=FALSE, rot.per=0, colors=brewer.pal(8, "Dark2")) 
   dev.off()
}

```


# Timezones 

```{r, gig.width=3, fig.height=2}
plt = ggplot(users_df, aes(x = reorder(timezone,timezone,
                     function(x)-length(x)))) + 
  geom_bar(stat = "count") + 
  theme_bw() + 
  # theme(axis.text.x = element_text(angle = 90,hjust=0.95,vjust=0.2)) + 
     theme(axis.text.x = element_text(angle = 45,hjust=1,vjust=1)) + 
  ylab("Number of Users") + 
  # scale_y_log10() +
  ggtitle("Users Timezones" )  + 
  theme(legend.position = "none") + 
  xlab("Timezones")
ggplotly(plt)
```


```{r, gig.width=3, fig.height=2}
ggsave('viz/timezones.pdf', width = 7, height = 4)
```


# Channel activity

```{r}


# tot = aggregate(df_channel_activity_monthly$activity, by = list(channel =df_channel_activity_monthly$channel), FUN = sum)

tb = table(df_texts$channel)
tot = data.frame('channel'=names(tb), x = as.numeric(tb))
# tot$x = tot$x/sum(tot$x)

 plt = ggplot(tot, aes(x=reorder(channel,x), y = x)) + 
   geom_bar(stat = "identity") + theme_bw(base_size = 15) + 
   # theme(axis.text.x = element_text(face = "bold", angle = 90)) + 
     theme(axis.text.x = element_text(angle = 45,hjust=1,vjust=1)) + 
   xlab("") + ylab("Number of Posts") + ggtitle("Total number of posts" )  + theme(legend.position = "none") + coord_flip() 
 #+ scale_y_log10()
 
ggplotly(plt)
```



```{r, gig.width=3, fig.height=2}
ggsave('viz/channels.pdf', width = 7, height = 8)
```


Country of Origin


Sankey Location

```{r}

country = read.csv("data_files/participants.csv", stringsAsFactors = FALSE)

nodes = data.frame(Name = union(country$Nationality, country$Team.Name))

links = country[,c("Team.Name", "Nationality")] %>% group_by(Team.Name, Nationality) %>% summarise(count = n())
links$IDsource <- match(links$Team.Name, nodes$Name)-1 
links$IDtarget <- match(links$Nationality, nodes$Name)-1

plt = sankeyNetwork(Links = links, Nodes = nodes,
                     Source = "IDsource", Target = "IDtarget",
                     Value = "count", NodeID = "Name", 
                     sinksRight=FALSE, fontSize = 13)
plt

visSave(plt, 'viz/sankey_nationalities.html')

#ggsave('viz/sankey_nationality.pdf', height=5, width=10)
```

Sankey Timezone

```{r}
org_team_part = users_df[!users_df$Team %in% c("Organiser", "Mentor", "Project Partner", "NO IDEA", "External Resource"),]
org_team_part = org_team_part[!grepl("Dropped Out", org_team_part$Team),]

#org_team_part$Team = as.character(org_team_part$Team)

nodes = data.frame(Name = union(org_team_part$Team, org_team_part$timezone))
nodes$Name = as.factor(nodes$Name)


links = org_team_part[,c("Team", "timezone")] %>% group_by(Team, timezone) %>% summarise(count = n())
links$IDsource <- match(links$Team, nodes$Name)-1 
links$IDtarget <- match(links$timezone, nodes$Name)-1

plt = sankeyNetwork(Links = links, Nodes = nodes,
                     Source = "IDsource", Target = "IDtarget",
                     Value = "count", NodeID = "Name", 
                     sinksRight=FALSE, fontSize = 13)
plt

visSave(plt, 'viz/sankey_timezones.html')

#ggsave('viz/sankey_timezones.pdf', height=5, width=10)

```



No. of Posts + Heatmap (Cut after week 6)

# Channel dynamics (Heatmap of Activity)

```{r}

store = function(a)
{
  if (is.integer(a) & length(a)==0)
    return(0)
  else
    return(a)
}

weeks = sort(unique(df_channel_activity_weekly$weeks))
channels = unique(df_channel_activity_weekly$channel)

channel_act_mat = matrix(0, nrow = length(channels), ncol = length(weeks))


for (i in 1:length(weeks))
{
  for (j in 1:length(channels))
  {
    channel_act_mat[j,i] = store(df_channel_activity_weekly$activity[df_channel_activity_weekly$channel %in% channels[j] & df_channel_activity_weekly$weeks == weeks[i]])
  }
  
}

colnames(channel_act_mat) = as.POSIXct(weeks, origin = "1970-01-01")
rownames(channel_act_mat) = channels


```



```{r}
data <- df_channel_activity_weekly
months <- sort(unique(data$weeks))
channels <- unique(data$channel)

M <- matrix(NA,length(channels), length(weeks))

for (i in 1:nrow(data)){
  M[match(data$channel[i], channels), match(data$weeks[i], weeks)] <- data$activity[i]
}

rownames(M) <- channels
colnames(M) <- paste('Week',1:ncol(M))

```


```{r, fig.height=5, fig.width=4}

pdf('viz/channels_heatmap.pdf', 7,10)

peak <- apply(M,1, which.max)
M1 <- M / apply(M,1,max, na.rm=T)

M1[is.na(M1)] <- 0
heatmap.0(M1[order(peak),1:6], cexRow = 1, cexCol = 2, col=colorRampPalette(c("white","darkred")), mar=c(10,15))

dev.off()
```


```{r, fig.width=3, fig.asp=1}

tot_posts <- apply(M,2,sum, na.rm=T)

pdf('viz/activity_total.pdf', 6,6)
plot.0(tot_posts, type='o', 
       ylab='Number of posts', xlab='Week', lwd=2)
dev.off()
```

Aggregated Mentions and Reaction Networks

```{r}
#Overall - mentions

text_ment1 = merge(df_text_mentions, users_df[,c("id", "Team")], by.x = "from", by.y = "id")
colnames(text_ment1)[colnames(text_ment1) == "Team"] = "From_Team"

text_ment1 = merge(text_ment1, users_df[,c("id", "Team")], by.x = "to", by.y = "id")
colnames(text_ment1)[colnames(text_ment1) == "Team"] = "To_Team"

g_grouped_mentions = graph_from_data_frame(text_ment1[,c("From_Team", "To_Team", "timestamp", "channel", "from", "to")], directed = TRUE)

E(g_grouped_mentions)$weight = 1

g_grouped_mentions = igraph::simplify(g_grouped_mentions, remove.loops = FALSE, remove.multiple = TRUE)

V(g_grouped_mentions)$print = V(g_grouped_mentions)$name

plt = visPlot_ni(g_grouped_mentions, edgewidth = rescale(E(g_grouped_mentions)$weight, to = c(1,10)), nodesize = degree(g_grouped_mentions, mode = "in"))

write.graph(g_grouped_mentions, "networks/grouped_mentions.graphml", format = "graphml")

plt

visSave(plt, 'networks/network_grouped_mention.html')
```

```{r}
#Overall - reactions

reac_ment1 = merge(df_emoticon_responses, users_df[,c("id", "Team")], by.x = "reaction_by", by.y = "id")
colnames(reac_ment1)[colnames(reac_ment1) == "Team"] = "From_Team"

reac_ment1 = merge(reac_ment1, users_df[,c("id", "Team")], by.x = "source_message_from", by.y = "id")
colnames(reac_ment1)[colnames(reac_ment1) == "Team"] = "To_Team"

g_grouped_reactions = graph_from_data_frame(reac_ment1[,c("From_Team", "To_Team", "source_message_timestamp", "channel", "source_message_from", "reaction_by")], directed = TRUE)

E(g_grouped_reactions)$weight = 1

g_grouped_reactions = igraph::simplify(g_grouped_reactions, remove.loops = FALSE, remove.multiple = TRUE)

V(g_grouped_reactions)$print = V(g_grouped_reactions)$name

plt = visPlot_ni(g_grouped_reactions, edgewidth = rescale(E(g_grouped_reactions)$weight, to = c(1,10)), nodesize = degree(g_grouped_reactions, mode = "in"))

write.graph(g_grouped_reactions, "networks/grouped_reactions.graphml", format = "graphml")

plt

visSave(plt, 'networks/network_grouped_reactions.html')

```


Channel wise interaction networks

```{r}

library(RColorBrewer)

#using all interactions

pdf('plotTeamNetworks.pdf')
par(mfrow=c(5,4), mar=c(0,0,2,0), cex.main=3)

df_emo = df_emoticon_responses[,c(1,2,3,5)]
colnames(df_emo) = c("From", "To", "Timestamp", "Channel")
df_emo$Type = "emoticon_responses"


df_tmt = df_text_mentions[!df_text_mentions$to %in% c("channel", "here", "everyone"),]
colnames(df_tmt) = c("From", "To", "Timestamp", "Channel")
df_tmt$Type = "text_mentions"

df_total = rbind(df_tmt, df_emo)


net_list = list()
j = 1

#channels = unique(df_total$Channel) 

for (i in teams)
{
  
    subs = df_total[df_total$Channel == i,]
  
    g_temp = graph_from_data_frame(subs, vertices = users_df[,c("id", "Team", "index", "comp_team")])
  
    write.graph(g_temp, paste("channel_nets/", i, ".graphml", sep = ""), format = "graphml")
  
    E(g_temp)$weight = 1
    g_simp_temp = igraph::simplify(g_temp, remove.multiple = TRUE, remove.loops =  FALSE)
  
    g_simp_temp = delete_vertices(g_simp_temp, v = V(g_simp_temp)[degree(g_simp_temp) == 0])
    
    V(g_simp_temp)$Name = V(g_simp_temp)$comp_team
  
    #print(col)
  
    net_list[[j]] = g_simp_temp
    
    j = j + 1
    
    plotNetwork(g_simp_temp, main = i, coms = V(g_simp_temp)$index)
  
}

```

Density order networks

```{r}


pdf('plotTeamNetworks_ordered.pdf')
par(mfrow=c(5,4), mar=c(0,0,2,0), cex.main=3)

ordered = order(sapply(net_list, edge_density))

for (i in ordered)
{
  g = net_list[[i]]
  
  plotNetwork(g, main = teams[i], coms = V(g)$index)
  
}

```

Other Channels

```{r}

pdf('plotChannelNetworks.pdf')
par(mfrow=c(2,2), mar=c(0,0,2,0), cex.main=3)

net_list = list()
j = 1

for (i in channels_df$name)
{
  if(!i %in% teams & nrow(df_total[df_total$Channel == i,]) > 0)
  {
    subs = df_total[df_total$Channel == i,]
  
    g_temp = graph_from_data_frame(subs, vertices = users_df[,c("id", "Team", "index", "comp_team")])
  
    write.graph(g_temp, paste("channel_nets/", i, ".graphml", sep = ""), format = "graphml")
  
    E(g_temp)$weight = 1
    g_simp_temp = igraph::simplify(g_temp, remove.multiple = TRUE, remove.loops =  FALSE)
  
    g_simp_temp = delete_vertices(g_simp_temp, v = V(g_simp_temp)[degree(g_simp_temp) == 0])
    
    V(g_simp_temp)$Name = V(g_simp_temp)$comp_team
  
    net_list[[j]] = g_simp_temp
    
    j = j + 1
    
    plotNetwork(g_simp_temp, main = i, coms = V(g_simp_temp)$index)
  }
}

```

Density order networks

```{r}


pdf('plotChannelNetworks_ordered.pdf')
par(mfrow=c(2,2), mar=c(0,0,2,0), cex.main=3)

ordered = order(sapply(net_list, edge_density))

for (i in ordered)
{
  g = net_list[[i]]
  
  plotNetwork(g, main = teams[i], coms = V(g)$index)
  
}

```


Interactions with Org. Team

```{r}

df_org_int = data.frame()

for (i in unique(df_total$Channel))
{
  subs = df_total[df_total$Channel == i,]
  
  part = users_df$id[!grepl("Dropped Out", users_df$Team)]
  
  n = nrow(subs[subs$To %in% users_df$id[users_df$Team %in% c("Mentor", "Organiser", "Project Partner")] & !subs$From %in% users_df$id[users_df$Team %in% c("Mentor", "Organiser", "Project Partner")],])
  
  n1 = nrow(subs[!subs$To %in% users_df$id[users_df$Team %in% c("Mentor", "Organiser", "Project Partner")] & subs$From %in% users_df$id[users_df$Team %in% c("Mentor", "Organiser", "Project Partner")],])
  
  df_org_int = rbind(df_org_int, data.frame(channel = i, messages_org = n+n1))
  
}

```

Plot

```{r}

plt = ggplot(df_org_int[df_org_int$channel %in% teams,], aes(x = reorder(channel, messages_org), y = messages_org)) + geom_bar(stat = "identity") + theme_bw(base_size = 20) + xlab("Teams") + ylab("# Interactions with Org Team") + ggtitle("Interactions with Organising Team") + theme(axis.text.x = element_text(angle = 0,hjust=1,vjust=1)) + coord_flip()

ggplotly(plt)

ggsave('viz/interactions_org_team.pdf', width = 8, height = 6)

```

*******************

Camille's Code

```{r}
library("gtsummary")

country %>%
  select(Gender, Age) %>%
  tbl_summary(
    statistic = list(all_continuous() ~ "{median} ({min}, {max})",
                     all_categorical() ~ "{n} ({p}%)"))

#ggsave('viz/gender_tbl_summary.pdf', width = 6, height = 5)
```

```{r}
ggplot(country)+
  aes(x = reorder(Nationality, Nationality, function(x) length(x))) +
  geom_bar(stat="count",width = 0.9, position = position_dodge())+
  scale_y_continuous(limits = c(0, 10), breaks = c(0,2,4,6,8,10))+
  labs(y = "Count",x = "Nationality")+
  theme_bw() + geom_text(stat = "count", aes(label = ""),
            position = position_dodge(width = 0.9), size = 2.5, hjust = -0.5)+
  coord_flip()

ggsave('viz/Nationality.pdf', width = 6, height = 5)
```  

```{r}

ggplot(country)+
  aes(x = reorder(Country.of.Residence, Country.of.Residence, function(x) length(x))) +
  geom_bar(stat="count",width = 0.9, position = position_dodge())+
  scale_y_continuous(limits = c(0, 10), breaks = c(0,2,4,6,8,10))+
  labs(y = "Count",x = "Country of origin")+
  theme_bw() + geom_text(stat = "count", aes(label = ""),
            position = position_dodge(width = 0.9), size = 2.5, hjust = -0.5)+
  coord_flip()

ggsave('viz/Country_of_origin.pdf', width = 6, height = 5)
```

Ecolution has one more member than the plot in the report. That is because of one user who listed their own name as the team name. I matched the user with their corresponding team - Ecolution. 

```{r}
ggplot(country)+
  aes(x = reorder(Team.Name, Team.Name, function(x) length(x))) +
  geom_bar(stat="count",width = 0.9, position = position_dodge())+
  scale_y_continuous(limits = c(0, 5), breaks = c(0,1,2,3,4, 5))+
  labs(y = "Count",x = "Teams")+
  theme_bw() + geom_text(stat = "count", aes(label = ""),
            position = position_dodge(width = 0.9), size = 2.5, hjust = -0.5)+
  coord_flip()

ggsave('viz/Team_size.pdf', width = 6, height = 5)
```

```{r}

country %>%
  ggplot( aes(x=Gender, y=Age, fill=Gender)) +
  geom_boxplot()+
  scale_fill_brewer()+
  geom_jitter(color="black", size=0.4, alpha=0.9)+
  scale_y_continuous(breaks = c(10, 15, 20, 25, 30, 35, 40, 45))+
  labs(x = "", y="Age", fill = "Gender")+
  theme_bw()

ggsave('viz/Gender_Age.pdf', width = 6, height = 5)
```

